The main imaging 'work horse' is the code ``LOFAR-Imag". It can be run with several different flavors that are described in \secref{Explore} till \secref{E_Intf}. The general structure of the  input is (as specified in a plain text file with extension .in)
\begin{linenumbers}
\resetlinenumber
\begin{verbatim}
&Parameters
p=v
p=v
&end
- - - - -- - - -- -
specific input lines
\end{verbatim}
\end{linenumbers}

where  \verb!p=v! stands for a list of parameters that are assigned a value (the so-called "namelist" input part). A list of the possible parameters with the first section where they are described is given in \tabref{LOFLI-namelist}. Some of these parameters apply to (practically) all run-options and are described here, many others are more specific and will be described in the appropriate section.

The namelist parameters that are used in most options are the following,

\begin{linenumbers}
\resetlinenumber
\begin{verbatim}
 &Parameters  RunOption= "xxxxx"
 OutFileLabel= "xx"
 AntennaRange= 100.  ! Maximum distance (from the core) for the range of the antennas (in [km]).
 SaturatedSamplesMax= 5     ! Maximum number of saturates time-samples per chunk of data
 Calibrations= "Calibrations202202071319.dat"     ! The antenna time calibration file. Not used when running on simulated data!
 SignFlp_SAI= 142092, 142093     ! Station-Antenna Identifiers for those where the sign of the signal should be reversed.
! PolFlp_SAI=  0     ! Station-Antenna Identifiers for those where the even-odd signals should be interchanged.
 BadAnt_SAI= 5095, 7093, 11089, 13088, 13089, 17084, 17085, 17094, 17095, 101085
  141083, 141086, 141087, 167094, 169082, 169090, 169094     ! Station-Antenna Identifiers for those that are malfunctioning.
 ExcludedStat= "RS305"     ! Mnemonics of the stations that should be excluded.
&end
\end{verbatim}
\end{linenumbers}
where more parameters from \tabref{LOFLI-namelist} may be added. All text on a line after an exclamation mark is considered comment and not used. The following lines are obtained from experience with other flashes, where:
\\\verb!RunOption="xxxxx"! specifies the particular flavor of the program that should be used, where "xxxxx" can be any of the following:
\begin{itemize}
\item \verb!RunOption="Explore"! for first exploration of this flash in order to get some idea of the layout and timing, see \secref{Explore}.
\item \verb!RunOption="Calibrate"! for performing time calibration using the Hilbert envelopes of the cross correlations, see \secref{Calibration}.
\item \verb!RunOption="ImpulsiveImager"! for running the impulsive Imager, see \secref{Imag}.
\item \verb!RunOption="FieldCalibrate"! Field Calibration for the TRI-D interferometric imager, see \secref{FCal}.
\item \verb!RunOption="TRI-D"! for the TRI-D imager with polarization observables, accounting for antenna function, see \secref{Intf}.
\item \verb!RunOption="ATRID"! for the ATRI-D imager with polarization observables, accounting for antenna function, see \secref{Intf}.
\item \verb!RunOption="SelectData"! to select real data, possibly for setup of simulation runs using program "SimulateData", see \secref{SelDat}.
\end{itemize}
\verb!OutFileLabel= "xx"! specifies an identifier for this particular run. It will be included in the name of the text-output file (with extension .out), as well as any figures (.pdf) and data files (.csv or .dat).
\\\verb!AntennaRange= 100.! Maximum distance (from the core, in [km]) for  antennas to be included in the calculations. Only in the "Explore" runoption this variable is not used.
\\\verb!SaturatedSamplesMax= 5! specifies the maximum number of time-samples in a data chunk (block of time trace) where the LOFAR digitizer has saturated. These are excluded from further processing. Experience teaches that =100 may be better since otherwise too many stations close to the flash will be excluded
\\\verb!Calibrations= "......"! points to a calibration file in the sub-folder \verb!Book! that was produces in an earlier calibration run, see \secref{Calibration} and/or  \secref{Intf}.
\\\verb!SignFlp_SAI=! list the antenna IDs where the signal is reversed.
\\\verb!PolFlp_SAI=! list the antenna IDs where the polarization is reversed.
\\\verb!BadAnt_SAI=! list the antenna IDs where the signal is bad.This line may be copied from the output of the RFI mitigation run, see \secref{RFI-out}.
\\\verb!ExcludedStat=! lists the stations that should be excluded from the calculations.
\\ Note that none of these lines in the namelist have a comma at the end, for the other input lines this is optional. Several keywords may appear on a single line if separated by a comma. Anything after an exclamation mark (within the namelist) is treated as a comment. A complete list of namelist keywords is given in \tabref{LOFLI-namelist}.


